Reproducibility of COVID-19 pre-prints

### 2021-07-22 outputs/paper/paper.Rmd
As submitted to arXiv

---
title: "Reproducibility of COVID-19 pre-prints in 2020"
subtitle: "Markers of open code or data not found in 81 per cent of pre-prints"
author: 
  - Annie Collins^[University of Toronto.]
thanks: "Collins thanks CANSSI Ontario for financial support. Code and data are available at: https://github.com/anniecollins/reproducibility."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "We create a dataset of all the pre-prints published on arXiv, bioRxiv, and medRxiv between 28 January 2020 and 31 January 2021. We extract the text from these pre-prints and parse them looking for keyword markers signalling the availability of the data and code underpinning the pre-print. We are unable to find markers of either open data or open code for 81 per cent of the pre-prints on medRxiv  in our sample, and X per cent of those on arXiv in our sample. Our paper demonstrates the need to have authors categorize the degree of openness of their pre-print as part of the pre-print submissions process, and more broadly, the need to better integrate open science training into a wide range of fields."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

@@

```{r loadpackages, include=FALSE}
library(here)
library(knitr)
library(lubridate)
library(tidyverse)
library(stargazer)
library(wesanderson)
```

```{r loaddata, include=FALSE}
med_open_data_results <- 
  read_csv(here::here("outputs/data/med_open_data_results.csv"))
med_results <- 
  read_csv(here::here("outputs/data/med_results.csv")) # All COVID medRxiv papers (9,929)

bio_open_data_results <- 
  read_csv(here::here("outputs/data/bio_open_data_results.csv"))
bio_results <- 
  read_csv(here::here("outputs/data/bio_results.csv")) # All COVID bioRxiv papers (3,031)

arxiv_open_data_results <- 
  read_csv(here::here("outputs/data/arxiv_open_data_results.csv"))
all_arxiv_results <- 
  read_csv(here::here("outputs/data/arxiv_results.csv"))

@@

Scientists use open repositories of papers to more quickly disseminate their research than is possible in traditional journals or conference proceedings. These repositories, such as arXiv, bioRxiv, medRxiv, and SocArXiv, are a critical component of science and many results build on the work published there. Thus it is important that the results that are published are credible. Pre-print repositories have drawn unprecedented attention in the context of the 2019 novel coronavirus (COVID-19) pandemic and the changes it has imposed on the scientific community [@else2020]. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a pre-print

Neither peer-review nor credentials are a panacea nor a guarantee of quality. And the gate-keeping and slow publication times of traditional journals mean pre-print repositories have an important role to play. But given the importance of these repositories, it is important that scientists impose on themselves various standards for their results. Following @weissgerber2021automated we examine papers about COVID-19 published to  arXiv, bioRxiv, medRxiv, and SocArXiv from January 2020, through to February 2021. We search for markers of open science as indicators of reproducibility, specifically open data and open code.

We find that approximately 81 per cent of sampled papers contain neither open data nor open code markers. Examining trends over time, we find that the proportion of pre-prints containing open data or code markers has fluctuated but shown no obvious increasing or decreasing trend throughout the pandemic. We also find that the presence of open data or open code markers has little influence on a pre-print's future publication, while the subset of sampled pre-prints that has been published contains an overall lower proportion of papers with these markers. Finally, we briefly examine the presence of open data and code markers in specific types of pre-prints, namely modelling, simulation, and machine learning research.

The remainder of this paper is structured as follows: in Section \@ref(data-and-methodology) we discuss the process of constructing our dataset through retrieving pre-prints from the medRxiv repository and mining them for open data and open code markers. In Sections \@ref(model) and \@ref(results) we explore some key findings in this raw data and explore some logistic regression models that support these findings. Finally, in Section \@ref(discussion) we discuss the implications of these findings in the broader context of reproducibility and science during the COVID-19 pandemic, as well as next steps to expand on our findings and questions raised in the research process.

# Data and Methodology

## Overview

Our primary dataset consists of information extracted from the arXiv, bioRxiv, medRxiv, and SocArXiv pre-print repositories. We downloaded pre-prints that contained various keywords related to COVID-19 (a full list of these is available in Appendix \@ref(keywords)), and then identified the availability of open code and open data using the Open Data Detection in Publications (ODDPub) text mining algorithm [@citePDDPub] in the `oddpub` package [@citeODDPubpackage].

We constructed this dataset by first creating a local copy of the repository using the repository API. We identified papers related to the COVID-19 pandemic through several keyword searches to create our dataset of papers and associated metadata. This metadata varies by repository, but generally includes the following information: title, abstract, author(s), date posted, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, and published DOI (if the pre-prints has since been published in a peer-reviewed journal). In the case of SocArXiv there are additionally fields that are self-reported by the submitting author about the availability of code and data.

We then checked these pre-prints for open data and code markers using the ODDPub algorithm. This required downloading each pre-print as a PDF and then converting the PDFs to text files. We then conducted the open data and open code detection procedure. This involved searching for keywords and other markers of open data and open code such as references to GitHub repositories and similar. A full listing of these markers is included in Appendix \@ref(markers). This was conducted using the `open_data_search()` function from the `oddpub` package.

The result of this process is a dataset indicating the presence of open data or open code markers in each pre-print (with a binary variable for each marker, and the relevant open data or open code statements when applicable). Our final analysis dataset was formed by joining these two datasets together, typically using the DOI or the unique filename, to form a dataset including all original, qualitative metadata information for each pre-print alongside its open data and open code status and markers.

@@

all <- rbind(just_arxiv, just_medRxiv, just_bioRxiv)

all %>% 
  mutate(month = floor_date(date, "month")) %>% 
  filter(year(date) >= 2020) %>% 
  ggplot(aes(x = month)) + 
  geom_histogram(aes(fill = type), stat="count", position = "dodge") + 
  labs(title = "Pre-prints posted per month",
       subtitle = "Total of 3,379 relevant pre-prints on arXiv, 3,031 on bioRxiv, 9,929 on medRxiv",
       x = "Month",
       y = "Number of pre-prints",
       fill = "Repository") +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Darjeeling1"))
```

## Summary statistics

The number of pre-prints posted per month increased dramatically between January and May 2020, reaching its maximum in May and subsequently decreasing. The number of pre-prints posted since August 2020 has remained reasonably steady at around 750 per month (Figure \@ref(fig:monthly-papers-total)). For context, COVID-19 was declared a pandemic by the World Health Organization (WHO) on March 11, 2020, at which point the number of cases globally had just surpassed 118,000 (primarily in east Asia) and the virus had been reported in 114 countries [@citeWHOtimeline].

Broadly, we are unable to find markers of either open data or open code for around 81 per cent of the 1,200 sampled pre-prints. The remaining 19 per cent of pre-prints was comprised of 4 per cent of the sample containing markers for open code, 8 per cent of the sample with markers for open data, and 7 per cent of the sample with markers for both open data and open code (Table \@ref(tab:summarycounts-medRxiv)).

```{r summarycounts-medRxiv}
# open data/open code count summary table
med_open_data_results %>% 
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>% 
  kable(col.names = c("Open data markers", "Open code markers", "Count", "Proportion of total"),
        caption = "Counts and proportions of open data and code markers in medRxiv sample",
        booktabs = TRUE,
        digits = 2)
```

```{r summarycounts-bioRxiv}
# open data/open code count summary table
bio_open_data_results %>% 
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>% 
  kable(col.names = c("Open data markers", "Open code markers", "Count", "Proportion of total"),
        caption = "Counts and proportions of open data and code markers in bioRxiv sample",
        booktabs = TRUE,
        digits = 2)
```

The distribution of total sampled pre-prints and sampled pre-prints with open data or code markers roughly follows that of COVID-19-related pre-prints posted in general (Figure \@ref(fig:monthly-papers-condition-stack)). The proportion of pre-prints with open data or code has fluctuated over time but shows no consistent overall increase or decrease throughout the course of the pandemic, nor in conjunction with increases or decreases in the total number of pre-prints posted to medRxiv (Figure \@ref(fig:monthly-proportion-of-open)). In our dataset, only one pre-print was sampled for the month of January 2020. This pre-print contained neither open data nor open code markers, thus the 0 per cent rate of open data and code for this month will be treated as an outlier, and removed modelling purposes.

```{r monthly-papers-condition-stack, fig.cap = "Number of pre-prints related to COVID-19 sampled from medRxiv, distinguished by presence of open data or code markers", out.width="90%", fig.align = "center"}
# Number of papers for each open data/code status per month
open_status <- 
  med_open_data_results %>% 
  select(title, date, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code)
open_status$condition[open_status$condition == 2] <- "Both"
open_status$condition[open_status$condition == 0] <- "Neither"
open_status$condition[open_status$condition == 1 & open_status$is_open_code == 1] <- "Open Code"
open_status$condition[open_status$condition == 1 & open_status$is_open_data == 1] <- "Open Data"

open_status %>% 
  ggplot() + 
  geom_bar(aes(x = factor(floor_date(date, "month")), 
               fill=factor(condition, levels = c("Neither", "Open Code", "Open Data", "Both")), 
               position = "stack")) +
  labs(fill = "Type of marker") +
  scale_x_discrete(breaks=c("2020-01-01","2020-04-01","2020-07-01", "2020-10-01", "2021-01-01"), 
                   labels=c("Jan 2020","Apr 2020","Jul 2020", "Oct 2020", "Jan 2021")) +
  ylab("Number of pre-prints") +
  xlab("Month") +
  labs(title="Sampled pre-prints by month", subtitle="medRxiv, Total = 1,200") +
  theme_minimal()
```

```{r monthly-proportion-of-open, fig.cap= "Proportion of pre-prints in sample with open data or open code markers by month", out.width="90%", fig.align = "center"}
# Proportion of papers posted with open data or open code each month
colors <- c("Open Data" = "blue", "Open Code" = "green")
med_month_sum_cov <- 
  med_open_data_results %>% 
  group_by(month = floor_date(date, "month")) %>% 
  summarise(count=n(), prop_open_code = sum(is_open_code)/n(), prop_open_data = sum(is_open_data)/n()) %>% 
  mutate(num_days = days_in_month(month),
         daily_rate = count/num_days)
med_month_sum_cov %>%
  ggplot() +
  geom_line(data = med_month_sum_cov, aes(month, prop_open_data, color = "Open data")) +
  geom_line(data = med_month_sum_cov, aes(month, prop_open_code, color = "Open code")) +
  ggtitle("Proportion of sampled pre-prints with open data or code markers") +
  labs(colour = "Type of marker",
       x = "Month",
       y = "Proportion of pre-prints") + 
  theme_minimal()
```

@@

In Table \@ref(tab:open-published-summary) we disaggregate pre-prints by whether there is an indication that it was published. We find that the proportion of pre-prints with open data or code markers among those that have been published is approximately five per cent lower compared with pre-prints that have not been published (Table \@ref(tab:open-published-summary)).

```{r open-published-summary}
published_open_status <- 
  med_open_data_results %>% 
  select(title, date, published, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code)
published_open_status$condition[published_open_status$condition == 2] <- "Both"
published_open_status$condition[published_open_status$condition == 0] <- "Neither"
published_open_status$condition[published_open_status$condition == 1 & published_open_status$is_open_code == 1] <- "Open code"
published_open_status$condition[published_open_status$condition == 1 & published_open_status$is_open_data == 1] <- "Open data"

published_open_status %>% 
  count(published, condition) %>% 
  pivot_wider(names_from = condition, values_from = n) %>%
  mutate(published = ifelse(published == 0, "No", "Yes"), `Proportion with either/both`= round((`Open data` + `Open code` + Both)/(`Open data` + `Open code` + Both + Neither), 2)) %>%
  rename(Published = published) %>%
  kable(caption = "Counts and proportions of open data markers by whether the pre-print was published",
        booktabs = TRUE)
```

There was a low number of papers that had markers of either open data or code in both those that were published and those that were not published (Figure \@ref(fig:open-published-stack)).

```{r open-published-stack, fig.cap= "Number of pre-prints in sample with open data or open code markers by publication status", out.width="90%", fig.align = "center"}
published_open_status$published[published_open_status$published == 0] <- "Published"
published_open_status$published[published_open_status$published == 1] <- "Unpublished"

>>>>

---
title: "Reproducibility of COVID-19 pre-prints"
author: 
  - Annie Collins^[University of Toronto.]
  - Rohan Alexander^[University of Toronto, rohan.alexander@utoronto.ca.]
thanks: "We thank Amy Farrow, Jessica Gronsbell, Monica Alexander, and Thomas William Rosenthal for helpful comments. Collins thanks CANSSI Ontario and Monica Alexander for financial support. Code and data are available at: https://github.com/anniecollins/reproducibility_markers_in_covid19_preprints."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "To examine the reproducibility of COVID-19 research, we create a dataset of pre-prints posted to arXiv, bioRxiv, medRxiv, and SocArXiv between 28 January 2020 and 30 June 2021 that are related to COVID-19. We extract the text from these pre-prints and parse them looking for keyword markers signalling the availability of the data and code underpinning the pre-print. For the pre-prints that are in our sample, we are unable to find markers of either open data or open code for 75 per cent of those on arXiv, 67 per cent of those on bioRxiv, 79 per cent of those on medRxiv, and 85 per cent of those on SocArXiv. We conclude that there may be value in having authors categorize the degree of openness of their pre-print as part of the pre-print submissions process, and more broadly, there is a need to better integrate open science training into a wide range of fields."
output:
  bookdown::pdf_document2:
    keep_tex:  true
toc: FALSE
bibliography: references.bib
---

@@

```{r loadpackages, include=FALSE}
library(here)
library(knitr)
library(lubridate)
library(tidyverse)
library(stargazer)
library(wesanderson)
library(kableExtra)
library(DiagrammeR)
```

```{r loaddata, include=FALSE}
med_open_data_results <- 
  read_csv(here::here("outputs/data/med_open_data_results.csv"))
med_results <- 
  read_csv(here::here("outputs/data/med_results.csv")) # All COVID medRxiv papers (12,060)
med_2019_open_data_results <- 
  read_csv(here::here("outputs/control-data/med_2019_open_data_results.csv")) # All pre-prints posted to medRxiv in 2019

bio_open_data_results <- 
  read_csv(here::here("outputs/data/bio_open_data_results.csv"))
bio_results <- 
  read_csv(here::here("outputs/data/bio_results.csv")) # All COVID bioRxiv papers (3,714)
bio_2019_open_data_results <- 
  read_csv(here::here("outputs/control-data/bio_2019_open_data_results.csv"))
bio_2019_results <- 
  read_csv(here::here("outputs/control-data/bio_data_2019.csv"))

arxiv_open_data_results <- 
  read_csv(here::here("outputs/data/arxiv_open_data_results.csv"))
all_arxiv_results <- 
  read_csv(here::here("outputs/data/arxiv_results.csv")) # All COVID arXiv papers (3,812)
arxiv_2019_open_data_results <- 
  read_csv(here::here("outputs/control-data/arxiv_2019_open_data_results.csv"))
all_arxiv_2019_results <- 
  read_csv(here::here("outputs/control-data/arxiv_data_2019.csv"))

socarxiv_open_data_results <- 
  read_csv(here::here("outputs/data/socarxiv_open_data_results.csv"))
socarxiv_2019_open_data_results <- 
  read_csv(here::here("outputs/control-data/socarxiv_2019_open_data_results.csv"))
socarxiv_metadata_2019 <- 
  read_csv(here::here("outputs/control-data/socarxiv_metadata_2019.csv"))

@@

Scientists use open repositories of papers to disseminate their research more quickly than is possible in traditional journals or conference proceedings. These repositories, such as arXiv, bioRxiv, medRxiv, and SocArXiv, are a critical component of scientific communication and a lot of research builds on the pre-prints posted there. Pre-print repositories have been especially important during the 2019 novel coronavirus (COVID-19) pandemic and the changes it has imposed on the scientific community [@else2020]. The centrality of pre-prints to science means that it is important that the results that are posted are credible. These repositories are not peer-reviewed, and, in general, anyone with appropriate academic credentials can submit a pre-print.

Neither peer-review nor credentials are a panacea nor a guarantee of quality. And the gate-keeping and slow publication times of traditional journals mean pre-print repositories are important. But it is important that scientists impose standards on themselves, and arguably repositories have a role to play here. Following @weissgerber2021automated, we examine pre-prints about COVID-19 posted to arXiv, bioRxiv, medRxiv, and SocArXiv from 28 January 2020 through to 30 June 2021. By way of background, each of these four repositories has a different focus: arXiv is general, bioRxiv focuses on biological sciences, medRxiv focuses on health sciences, and SocArXiv focuses on social sciences. We search for markers of open science as indicators of reproducibility, specifically open data and open code.

We find that of the papers sampled, approximately 75 per cent of papers from arXiv, 67 per cent of papers from bioRxiv, 79 per cent of papers from medRxiv, and 85 per cent of papers from SocArXiv contain neither open data nor open code markers. A summary of our main results is contained in Figure \@ref(fig:flowchart). Examining trends over time, we find that the proportion of pre-prints containing open data or code markers has fluctuated but shown no obvious trend throughout the pandemic. We also find that the presence of open data or open code markers seems to have little association with a pre-print's subsequent publication, and the subset of sampled pre-prints that have been published contains approximately the same proportion of papers with these markers. 
<!-- Finally, we briefly examine the presence of open data and code markers in specific types of pre-prints, namely modelling, simulation, and machine learning research. -->

```{r flowchart, fig.cap = "Summary of process and main results", out.width="90%", fig.align = "center"}
DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = TB]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle]

all [label =  'All pre-prints posted between 28 January 2020 and 30 June 2021']
onlycovid [label =  'Those about COVID-19']
arxivcovidcount [label =  'arXiv: 4,250 pre-prints']
biorxivcovidcount [label =  'bioRxiv: 4,146 pre-prints']
medrxivcovidcount [label =  'medRxiv: 13,194 pre-prints']
socarxivcovidcount [label =  'SocArXiv: 478 pre-prints']
randomsample [label =  'Random sampling stratified by repository']
arxivsamplecount [label =  'arXiv: 1,000 pre-prints']
biorxivsamplecount [label =  'bioRxiv: 1,000 pre-prints']
medrxivsamplecount [label =  'medRxiv: 1,500 pre-prints']
socarxivsamplecount [label =  'SocArXiv: 478 pre-prints']
opencodeanddata [label =  'Those with markers of open code or open data']
arxivopencount [label =  'arXiv: 248 pre-prints']
biorxivopencount [label =  'bioRxiv: 326 pre-prints']
medrxivopencount [label =  'medRxiv: 320 pre-prints']
socarxivopencount [label =  'SocArXiv: 73 pre-prints']

# edge definitions with the node IDs
all -> onlycovid -> {arxivcovidcount biorxivcovidcount medrxivcovidcount socarxivcovidcount} -> randomsample -> {arxivsamplecount biorxivsamplecount medrxivsamplecount socarxivsamplecount} -> opencodeanddata -> {arxivopencount biorxivopencount medrxivopencount socarxivopencount}
}")
```

The remainder of this paper is structured as follows: in Section \@ref(methodology) we discuss the process of constructing our dataset through retrieving pre-prints from the arXiv, bioRxiv, medRxiv, and SocArXiv repositories and mining them for open data and open code markers. In Section \@ref(results), we present the results and key findings of this process. Finally, in Section \@ref(discussion) we discuss the implications of these findings in the broader context of reproducibility and science during the COVID-19 pandemic, as well as next steps to expand on our findings and questions raised in the research process.

# Methodology

## Pre-print metadata

Our primary dataset consists of pre-print metadata extracted from the arXiv, bioRxiv, medRxiv, and SocArXiv pre-print repositories via their respective Application Programming Interfaces (APIs). This metadata varies by repository, but generally includes the following information: title, abstract, author(s), date created, research field, DOI, version number, corresponding author, corresponding author's institutional affiliation, published DOI (if the pre-print has since been published in a peer-reviewed journal), and download link. The data collection process was conducted separately for COVID-19 and pre-COVID-19 papers.

For COVID-19-related pre-prints, we first created a local copy of each repository containing all metadata for pre-prints posted between 1 January, 2020, and 30 June, 2021. In the case of SocArXiv there are fields that are self-reported by the submitting author about the availability of code and data. We classified individual pre-prints as "COVID-19-related" based on whether they contained one or more of the following terms in their title or abstract (case insensitive): "COVID-19", "COVID 19", "corona virus", "coronavirus", "coronavirus-2", "SARS-CoV-2", "SARSCoV-2", and "2019-nCoV". We then randomly sampled pre-prints for further analysis, except for SocArXiv, for which we analyzed all pre-prints that matched the above criteria, because there were only 478.

For pre-COVID-19 pre-prints, we created a local copy of each repository containing all metadata for pre-prints posted between 1 January, 2019, and December 31, 2019. Since medRxiv was launched in June 2019, we used all pre-print data from the latter half of 2019. We then randomly sampled 1,200 pre-prints from each repository's dataset for analysis, except for medRxiv for which only 913 pre-prints were available over this time.

## Open data and code detection

We checked our sampled pre-prints for open data and code markers using the Open Data Detection in Publications (ODDPub) text mining algorithm [@citePDDPub] within the `oddpub` R package [@citeODDPubpackage]. This required downloading each pre-print as a PDF, which we did using `heapsofpapers` [@citealexandermafouz] and then converting the PDFs to text files. Some papers from SocArXiv were only available in Word document format (.docx) and were downloaded as such and then converted to text format. We then conducted the open data and open code detection procedure, which involved searching for keywords and other markers of open data and open code availability. This was conducted using the `open_data_search()` function from the `oddpub` package. What constitutes open code or open data is complicated and discipline specific. The details of the `oddpub` approach are available in @citePDDPub. The general criteria are that specific mention should be made of where the data and code are located, and that data should be as close to raw as possible. Our work was conducted using the statistical programming language R [@citeR].

The result of this process is a dataset indicating the presence of open data or open code markers in each pre-print (with a logical vector for each marker, followed by the relevant open data or open code statements where applicable). Our final dataset was formed by joining this output with the original sample metadata, typically using the DOI or the unique file name, to form a dataset including all original metadata for each pre-print alongside its open data and open code status and markers.

Although SocArXiv's metadata contained author-reported information concerning data and code availability for each pre-print, we elected to use the results of the ODDPub algorithm in our analysis to ensure comparability of results between pre-prints from all repositories. 
<!-- Notable discrepancies between the ODDPub algorithm and authors' self-reported data and code availability are documented in Appendix \@ref(socarxivissues). -->

@@

just_socarxiv <- 
  socarxiv_open_data_results  %>% 
  select(date_created) %>%
  rename(date = "date_created") %>%
  mutate(date = as.Date(date), type = "SocArXiv")

all <- rbind(just_arxiv, just_medRxiv, just_bioRxiv, just_socarxiv)

all %>% 
  mutate(month = floor_date(date, "month")) %>% 
  filter(year(date) >= 2020) %>%
  filter(date < "2021-06-30") %>% 
  ggplot(aes(x = month)) + 
  geom_histogram(aes(fill = type), stat="count", position = "dodge") + 
  labs(title = "COVID-19-related pre-prints posted per month",
       subtitle = "Total of 4,250 relevant pre-prints on arXiv, 4,146 on bioRxiv, 13,194 on medRxiv, and \n478 on SocArXiv.",
       x = "Month",
       y = "Number of pre-prints",
       fill = "Repository") +
  scale_x_date(breaks=as.Date(c("2020-01-01","2020-07-01", "2021-01-01")),
                   labels=c("Jan 2020", "Jul 2020", "Jan 2021")) +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Darjeeling1")) +
  facet_wrap(~type,
             scales = "free_y")
```

## All pre-prints related to COVID-19

The number of pre-prints posted per month increased in the first half of 2020 across all repositories, reaching a maximum sometime between April and June (depending on repository) and subsequently decreasing. The number of pre-prints posted monthly since August 2020 has remained reasonably steady, with the exception of medRxiv, which experienced an increase to nearly 1,000 pre-prints posted in March 2021 (Figure \@ref(fig:monthly-papers-total)). For context, COVID-19 was declared a pandemic by the World Health Organization (WHO) on March 11, 2020, at which point the number of cases globally had just surpassed 118,000 (primarily in east Asia) and the virus had been reported in 114 countries [@citeWHOtimeline].

## Open data and code

From the collection of all pre-prints related to COVID-19, we randomly sampled 3,910 pre-prints to analyze, stratified by repository. This sample is broken down as follows: 1,500 from medRxiv, 1,000 from arXiv, 1,000 from bioRxiv, and 478 from SocArXiv. Broadly, we are unable to find markers of either open data or open code for 3,011 pre-prints or approximately 76 per cent of our sample (Table \@ref(tab:summarycounts-repositories)). Of the remaining pre-prints, 7 per cent contained open code markers only, 10 per cent contained open data markers only, and 8 per cent included markers of both open data and open code.

When differentiated by repository, we observe that open data and code markers were absent from 75 per cent of the sampled arXiv pre-prints, 67 per cent of the sampled bioRxiv pre-prints, 79 per cent of the sampled medRxiv pre-prints, and 85 per cent of all COVID-19-related SocArXiv pre-prints. The distribution of the remaining portion of pre-prints also varies by repository (Table \@ref(tab:summarycounts-repositories)). Notably, 28 per cent of sampled pre-prints from bioRxiv contained open data markers and 22 per cent of sampled arXiv pre-prints contained markers of open code, the highest proportions of any repository for each type of marker. Our results are similar to @mcguinness2021descriptive, who focus on medRxiv and find that 23 per cent describe open data.

```{r, summarycounts-overall}
just_arxiv <-
  arxiv_open_data_results  %>%
  select(submitted, is_open_data, is_open_code) %>%
  mutate(submitted = as.Date(submitted)) %>%
  mutate(type = "arXiv") %>%
  rename(date = submitted)

just_medRxiv <-
  med_open_data_results  %>%
  mutate(date = as.Date(date))  %>%
  select(date, is_open_data, is_open_code) %>%
  mutate(type = "medRxiv")

just_bioRxiv <-
  bio_open_data_results  %>%
  select(date, is_open_data, is_open_code) %>%
  mutate(date = as.Date(date)) %>%
  mutate(type = "bioRxiv")

just_socarxiv <-
  socarxiv_open_data_results  %>%
  select(date_created, is_open_data, is_open_code) %>%
  rename(date = "date_created") %>%
  mutate(type = "SocArXiv")

all <- rbind(just_arxiv, just_medRxiv, just_bioRxiv, just_socarxiv)

all_total <- 
  all %>%
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>%
  select(is_open_data, is_open_code, n, prop_total)
```

```{r summarycounts-repositories}
# open data/open code count summary table

all_by_type <- 
  all %>%
  group_by(type) %>%
  count(is_open_data, is_open_code) %>%
  mutate(prop_total = n / sum(n)) %>%
  mutate(is_open_data = ifelse(is_open_data == 0, "No", "Yes"),
         is_open_code = ifelse(is_open_code == 0, "No", "Yes")) %>%
  ungroup() %>%
  select(is_open_data, is_open_code, n, prop_total)

rbind(all_total, all_by_type) %>% 
  kable(col.names = c("Open data markers", "Open code markers", "Count", "Proportion of total"),
        caption = "Count and proportions of open data and code markers by pre-print repository",
        booktabs = TRUE,
        digits = 2,
        format.args = list(big.mark = ",")) %>%
  group_rows(index = c("Total" = 4, "arXiv" = 4, "bioRxiv" = 4, "medRxiv" = 4, "SocArXiv" = 4))
```

The distribution of total sampled pre-prints and sampled pre-prints with open data or code markers roughly follows that of COVID-19-related pre-prints posted in general (Figure \@ref(fig:monthly-papers-condition-stack)). The proportion of pre-prints with open data or code has fluctuated over time but shows no consistent overall increase or decrease throughout the course of the pandemic, nor in conjunction with increases or decreases in the total number of pre-prints posted to any given repository. In our datasets, very few (if any) pre-prints were sampled for the month of January 2020. None of these pre-prints contained open data or open code markers, thus the 0 per cent rate of open data and code for this month across all repositories should be considered an outlier.

```{r monthly-papers-condition-stack, fig.cap = "Number of sampled pre-prints related to COVID-19, distinguished by presence of open data or code markers", out.width="90%", fig.align = "center"}
# Number of papers for each open data/code status per month
open_status_med <- 
  med_open_data_results %>% 
  select(title, date, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "medRxiv")

open_status_bio <- 
  bio_open_data_results %>% 
  select(title, date, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "bioRxiv", date = as.character(date))

open_status_arxiv <- 
  arxiv_open_data_results %>% 
  select(title, submitted, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "arXiv", date = as.character(submitted) %>% str_sub(end = 10)) %>%
  select(title, date, is_open_code, is_open_data, condition, repo)

open_status_socarxiv <- 
  socarxiv_open_data_results %>% 
  select(title, date_created, is_open_code, is_open_data) %>% 
  mutate(condition = is_open_data + is_open_code, repo = "SocArXiv", date = as.character(date_created)) %>%
  select(title, date, is_open_code, is_open_data, condition, repo)

open_status <- rbind(open_status_med, open_status_bio, open_status_arxiv, open_status_socarxiv)

open_status$date <- as.Date(open_status$date)
open_status$condition[open_status$condition == 2] <- "Both"
open_status$condition[open_status$condition == 0] <- "Neither"
open_status$condition[open_status$condition == 1 & open_status$is_open_code == 1] <- "Open Code"
open_status$condition[open_status$condition == 1 & open_status$is_open_data == 1] <- "Open Data"

open_status %>% 
  ggplot() + 
  geom_bar(aes(x = factor(floor_date(date, "month")), 
              fill=factor(condition, levels = c("Neither", "Open Code", "Open Data", "Both")), 
              position = "stack")) +
  labs(fill = "Type of marker") +
  scale_x_discrete(breaks=c("2020-01-01","2020-07-01", "2021-01-01"), 
                   labels=c("Jan 2020", "Jul 2020", "Jan 2021")) +
  ylab("Number of pre-prints") +
  xlab("Month") +
  labs(title="Sampled pre-prints by month", subtitle="Sample size from arXiv is 1,000, from bioRxiv is 1,000, from medRxiv is 1,500, and from \nSocArXiv is 410. Note the different y-axes.") +
  theme_minimal() +
  scale_fill_manual(values = wes_palette("Darjeeling1")) +
  facet_wrap(~repo,
             scales = "free_y")
```

```{r monthly-proportion-of-open, fig.cap= "Proportion of pre-prints in sample with open data or open code markers by month", out.width="90%", fig.align = "center"}
# Proportion of papers posted with open data or open code each month
med_month_sum_cov <- 
  med_open_data_results %>% 
  mutate(date = as.Date(date)) %>%
  group_by(month = floor_date(date, "month")) %>% 
  summarise(count = n(), "Open Code" = mean(is_open_code), "Open Data" = mean(is_open_data)) %>% 
  mutate(daily_rate = count/days_in_month(month), repo = "medRxiv")

bio_month_sum_cov <- 
  bio_open_data_results %>% 
  mutate(date = as.Date(date)) %>%
  group_by(month = floor_date(date, "month")) %>% 
  summarise(count = n(), "Open Code" = mean(is_open_code), "Open Data" = mean(is_open_data)) %>% 
  mutate(daily_rate = count/days_in_month(month), repo = "bioRxiv")

arxiv_month_sum_cov <- 
  arxiv_open_data_results %>% 
  group_by(month = floor_date(submitted, "month")) %>% 
  summarise(count = n(), "Open Code" = mean(is_open_code), "Open Data" = mean(is_open_data)) %>% 
  mutate(daily_rate = count/days_in_month(month), repo = "arXiv")

socarxiv_month_sum_cov <- 
  socarxiv_open_data_results %>% 
  group_by(month = floor_date(as.Date(date_created), "month")) %>% 
  summarise(count = n(), "Open Code" = mean(is_open_code), "Open Data" = mean(is_open_data)) %>% 
  mutate(daily_rate = count/days_in_month(month), repo = "SocArXiv")

month_sum_cov <- rbind(med_month_sum_cov, bio_month_sum_cov, arxiv_month_sum_cov, socarxiv_month_sum_cov) %>%
  pivot_longer(cols = c("Open Data", "Open Code")) %>%
  filter(value != 1) # Removes outliers where small number of papers in a month was sampled and both had open data/code

# TODO: Fix date labels, facet_wrap labels
# month_sum_cov %>%
#   ggplot() +
#   geom_line(data = month_sum_cov, aes(month, value, color = repo)) +
#   ggtitle("Proportion of sampled pre-prints with open data or code markers") +
#   labs(colour = "Repository",
#        x = "Month",
#        y = "Proportion of pre-prints") + 
#   theme_minimal() +
#   scale_color_manual(values = wes_palette("Darjeeling1")) +
#   facet_wrap(~name)
```

@@

Our metadata also contains an indication of whether a pre-print has been published, in the form of a DOI linking to the corresponding published paper. The proportion of pre-prints that have been published varies by repository (Table \@ref(tab:published-sample-summary)). Notably, of all COVID-19-related pre-prints in our dataset, approximately 30 per cent of those posted to bioRxiv and nearly one third of those posted to medRxiv were published. This is high in comparison to the proportions from arXiv and SocArXiv, and although this might suggest that COVID-19-related pre-prints in biomedical fields have received greater attention overall than pre-prints from other fields, our results in Section \@ref(pre-pandemic-pre-prints) suggest that this pattern pre-dates the pandemic.

```{r published-sample-summary}
med_published <- 
  med_results %>% 
  select(title, date, published) %>% 
  mutate(repo = "medRxiv", published = as.numeric(!is.na(published)))

bio_published <- 
  bio_results %>% 
  select(title, date, published) %>% 
  mutate(repo = "bioRxiv", published = as.numeric(!is.na(published)))

arxiv_published <- 
  all_arxiv_results %>% 
  select(title, submitted, doi) %>%
  mutate(doi = as.numeric(!is.na(doi)), repo = "arXiv") %>%
  rename(published = doi, date = submitted)

socarxiv_published <- 
  socarxiv_open_data_results %>% 
  select(title, date_created, links.doi) %>%
  mutate(links.doi = as.numeric(!is.na(links.doi)), repo = "SocArXiv") %>%
  rename(published = links.doi, date = date_created)

published <- rbind(med_published, bio_published, arxiv_published, socarxiv_published)

published %>% 
  group_by(repo) %>% 
  summarise(`Published pre-prints` = sum(published), 
            `Proportion published` = mean(published) %>% round(2)) %>%
  kable(caption = "Counts and proportion of published COVID-19 pre-prints in each repository",
        col.names = c("Repository", "Published pre-prints", "Proportion published"),
        format.args = list(big.mark = ","))
```
